---
title: "AJIVE Analysis"
author: "Taebin Kim"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/taebinkim/Desktop/UNC/STOR 765/STOR765_Spring2023_TeamDS")
getwd()
```

```{r, warning=FALSE, message=FALSE, results='hide'}
source_rmd = function(file, ...) {
  tmp_file = tempfile(fileext = ".R")
  on.exit(unlink(tmp_file), add = TRUE)
  knitr::purl(file, output = tmp_file)
  source(file = tmp_file, ...)
}

source_rmd("00_manage_packages.Rmd")
```
## Setup

Install the `r_jive` package.
```{r}
devtools::install_github("taebinkim7/r_jive")
library(ajive)
```

Load preprocessed data blocks.
```{r}
load("data/cortisol.RData")
load("data/fmri.RData")
load("data/working.RData")
load("data/demographics2.RData")
```


Set signal ranks based on the PC scree plots.
```{r}
cortisol_signal_rank <- 3
fmri_signal_rank <- 5
working_signal_rank <- 9
demographics_signal_rank <- 4
```

Redefine data blocks for the AJIVE analyses.
```{r}
cortisol = cortisol_wide
fmri = fmri_wide
working = working_wide %>%
  rownames_to_column("SubjectID")
demographics = demographics_block2 %>%
  fastDummies::dummy_cols("Sex") %>%
  select(-c(Sex, Sex_F, AgeInYears))
```

Define a function for fitting an AJIVE model.

* Usage
```{r, eval=FALSE}
fit_ajive(data_blocks, signal_ranks, n_wedin_samples = 100, n_rand_dir_samples = 100)
```

* Arguments

`data_blocks`: List of data blocks for AJIVE analysis

`signal_ranks`: Predetermined signal ranks for the data blocks. The length should be the same as the number of data blocks.

`n_wedin_samples`: Number of samples for obtaining the Wedin bound. 

`n_rand_dir_samples`: Number of random direction samples

```{r}
fit_ajive <- function(data_blocks,
                      signal_ranks,
                      n_wedin_samples = 100,
                      n_rand_dir_samples = 100) {
  # keep overlapping subjects across data blocks
  merged_data_block <- Reduce(
    function (...) { merge(..., all = FALSE) }, data_blocks)
  common <- merged_data_block$SubjectID
  inter_data_blocks <- sapply(
    data_blocks,
    function(x) as.matrix(x[x$SubjectID %in% common, ] %>% select(-1))
  )
  
  # fit AJIVE
  ajive <- ajive::ajive(inter_data_blocks,
                        signal_ranks,
                        n_wedin_samples = n_wedin_samples,
                        n_rand_dir_samples = n_rand_dir_samples)
  return(ajive)
}
```

## AJIVE Analyses

* Q1. Investigate relationship between stress response profile (cortisol timeseries block) and brain region response (FMRI block) during stress test.
```{r}
data_blocks_cortisol_fmri <- list(cortisol, fmri)
ajive_cortisol_fmri <- fit_ajive(
  data_blocks_cortisol_fmri,
  c(cortisol_signal_rank, fmri_signal_rank)
)
cat("Joint rank:", ajive_cortisol_fmri$joint_rank)
```

* Q2. Investigate relationship between stress response profile (cortisol timeseries block) and brain region response (working memory block) during working memory task.
```{r}
data_blocks_cortisol_working <- list(cortisol, working)
ajive_cortisol_working <- fit_ajive(
  data_blocks_cortisol_working,
  c(cortisol_signal_rank, working_signal_rank)
)
cat("Joint rank:", ajive_cortisol_working$joint_rank)
```

* Q3. Repeat Q1 and Q2 with demographics block as the third block.
```{r, eval=FALSE}
data_blocks_cortisol_fmri_demographics <- list(cortisol, fmri, demographics)
ajive_cortisol_fmri_demographics <- fit_ajive(
  data_blocks_cortisol_fmri_demographics,
  c(cortisol_signal_rank, fmri_signal_rank, demographics_signal_rank)
)
cat("Joint rank:", ajive_cortisol_fmri_demographics$joint_rank)

data_blocks_cortisol_working_demographics <- list(cortisol,
                                                  working,
                                                  demographics)
ajive_cortisol_working_demographics <- fit_ajive(
  data_blocks_cortisol_working_demographics,
  c(cortisol_signal_rank, working_signal_rank, demographics_signal_rank)
)
cat("Joint rank:", ajive_cortisol_working_demographics$joint_rank)
```







