---
title: "Make Data"
author: "Kyungjin Sohn"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, warning=FALSE, message=FALSE, results='hide'}
source_rmd = function(file, ...) {
  tmp_file = tempfile(fileext = ".R")
  on.exit(unlink(tmp_file), add = TRUE)
  knitr::purl(file, output = tmp_file)
  source(file = tmp_file, ...)
}

source_rmd("00_manage_packages.Rmd")
```

## 1. Demographics data
```{r read_1}
demographics_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                                 sheet = "Demographics_block5")
```

We examine the structure of the demographics block :
```{r structure_1}
str(demographics_block)
```
We see that the demographics block contains 130 rows and 6 columns: 

1. _Subject_: subject identifier, which is a string, e.g. `cnt_101_bl`
2. _Sex_: gender of the subject, e.g. M(Male) or F(Female)
3. _AgeInMonths_: age of the subject in months
4. _PubertyTannerSelf_: subject's self puberty level
5. _PubertyTannerParent_: parent's puberty level
6. _ParentalEducationAvg_ : average education level of the subject's parent

We check for missing values:
```{r missings_1}
sapply(demographics_block, function(x) sum(is.na(x)))
```

* There are no missing values in _Subject_, _Sex_, and _AgeInMonths_.
* There are at least one missing value in _PubertyTannerSelf_, _PubertyTannerParent_, and _ParentalEducationAvg_.

### 1) Subject
No subject identifiers are duplicated: 
```{r dupes_1}
table(duplicated(demographics_block$Subject))
```

### 2) Sex
Next, we changed the Sex variable to a factor type variable.
```{r factor_1}
demographics_block <- demographics_block %>% 
  mutate(Sex = factor(Sex, levels = c("M", "F")))
str(demographics_block$Sex)
```

### 3) Age In Months
Since we only had the age in months, we added a new column to show the age in years rounded to the second decimal place.
```{r AgeInYears_1}
demographics_block <- demographics_block %>% 
     mutate(AgeInYears = round(AgeInMonths/12, 2)) %>% 
     relocate(AgeInYears, .after = AgeInMonths)
```

### 4 ~ 6) Own puberty level, Parent's puberty level, and Average parent education
The own puberty level variable ranges from 1 to 5, with 29 missing values. The parent's puberty level variable also ranges from 1 to 5, with 21 missing values. Lastly, the average parent education variable ranges from 1 to 9, with 6 missing values.
```{r summary_1}
summary(demographics_block)
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save_1}
save(demographics_block, file = "./data/demographics.RData")
```

## 2. Cortisol data
```{r read_2}
cortisol_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                              sheet = "Cortisol_MRI_block1")
```

We examine the structure of the cortisol block: 
```{r structure_2}
str(cortisol_block)
```
We see that the cortisol block contains 118 rows and 7 columns: 

1. _StudyID_: a patient identifier, which is a string, e.g. `cnt_101_bl`
2. _M_S1_, _M_S2_, ..., _M_S6_: the cortisol measure at each of six timepoints, which is a nonnegative number

We check for missing values:
```{r missings_2}
sapply(cortisol_block, function(x) sum(is.na(x)))
```

* There are no missing values.

### 1) StudyID
No subject identifiers are duplicated: 
```{r dupes_2}
table(duplicated(cortisol_block$StudyID))
```

### 2) M_S1 ~ M_S6
Since the cortisol data are time-series, we will want to pivot the six timepoints into a single column for EDA.
```{r pivot_2}
cortisol_long <- cortisol_block %>%
  pivot_longer(!StudyID, names_to="timepoint", values_to="cortisol") %>%
  mutate(timepoint = strtoi(str_sub(timepoint, 4)))   # extract timepoint and cast to integer
head(cortisol_long)
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save_2}
save(cortisol_block, cortisol_long, file = "./data/cortisol.RData")
```

## 3. FMRI data

```{r read_3}
fmri_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                         sheet = "ROI_MIST_block2")
```

We examine the structure of the FMRI block: 
```{r structure_3}
str(fmri_block)
```
We see that the fmri_block block contains 928 rows and 3 columns: 

1. _Subject_: a patient identifier, which is a string, e.g. `cnt_101_bl`
2. _ROI_: Region of Interest
3. _MeanPerSigChg_: Mean per signal change

We check for missing values:
```{r missings_3}
sapply(fmri_block, function(x) sum(is.na(x)))
```

* There are no missing values.

### 1) Subject
There are total of 116 unique subjects.
```{r unique_3_1}
length(unique(fmri_block$Subject))
```

### 2) ROI
There are total of 8 unique ROIs.
```{r unique_3_2}
length(unique(fmri_block$ROI))
```

### 3) MeanPerSigChg

Since there are 116 unique subjects and 8 unique ROIs, we will need to reshape the dataset from long to wide format. The value in each cell of the resulting dataset will be the _MeanPerSigChg_ for a given subject and ROI combination. This transformation will allow us to more easily compare the signal changes across different brain regions and subjects.
```{r pivot_3}
fmri_wide <- fmri_block %>% 
  pivot_wider(names_from = ROI, values_from = MeanPerSigChg)
head(fmri_wide, 5)
```
After applying this transformation, the resulting dataset have 116 rows and 9 columns, where the first column is the subject ID and the remaining 8 columns represent the ROIs.

After making the modifications mentioned above, we saved the updated data frame.
```{r save_3}
save(fmri_block, fmri_wide, file = "./data/fmri.RData")
```

## 4. Psych data

```{r}

```

## 5. Working memory data
## 6. Autonomic data
