---
title: "Make Data"
author: "Kyungjin Sohn"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r, warning=FALSE, message=FALSE, results='hide'}
source_rmd = function(file, ...) {
  tmp_file = tempfile(fileext = ".R")
  on.exit(unlink(tmp_file), add = TRUE)
  knitr::purl(file, output = tmp_file)
  source(file = tmp_file, ...)
}

source_rmd("00_manage_packages.Rmd")
```

## 1. Demographics data
```{r read_1}
demographics_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                                 sheet = "Demographics_block5")
```

We examine the structure of the demographics block :
```{r structure_1}
str(demographics_block)
```
We see that the demographics block contains 130 rows and 6 columns: 

1. **`Subject`**: subject identifier, which is a string, e.g. `cnt_101_bl`
2. **`Sex`**: gender of the subject, e.g. M(Male) or F(Female)
3. **`AgeInMonths`**: age of the subject in months
4. **`PubertyTannerSelf`**: subject's self puberty level
5. **`PubertyTannerParent`**: parent's puberty level
6. **`ParentalEducationAvg`**: average education level of the subject's parent

We check for missing values:
```{r missings_1}
sapply(demographics_block, function(x) sum(is.na(x)))
```

* There are no missing values in _Subject_, _Sex_, and _AgeInMonths_.
* There are at least one missing value in _PubertyTannerSelf_, _PubertyTannerParent_, and _ParentalEducationAvg_.

### 1) SubjectID
No subject identifiers are duplicated: 
```{r dupes_1}
demographics_block <- demographics_block %>% 
  rename("SubjectID" = "Subject")

table(duplicated(demographics_block$SubjectID))
```

### 2) Sex
Next, we changed the Sex variable to a factor type variable.
```{r factor_1}
demographics_block <- demographics_block %>% 
  mutate(Sex = factor(Sex, levels = c("M", "F")))
str(demographics_block$Sex)
```

### 3) Age In Months
Since we only had the age in months, we added a new column to show the age in years rounded to the second decimal place.
```{r AgeInYears_1}
demographics_block <- demographics_block %>% 
     mutate(AgeInYears = round(AgeInMonths/12, 2)) %>% 
     relocate(AgeInYears, .after = AgeInMonths)
```

### 4 ~ 6) Own puberty level, Parent's puberty level, and Average parent education
The own puberty level variable ranges from 1 to 5, with 29 missing values. The parent's puberty level variable also ranges from 1 to 5, with 21 missing values. Lastly, the average parent education variable ranges from 1 to 9, with 6 missing values.
```{r summary_1}
summary(demographics_block)
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save_1}
# save(demographics_block, file = "./data/demographics.RData")
```

## 2. Cortisol data
```{r read_2}
cortisol_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                              sheet = "Cortisol_MRI_block1")
```

We examine the structure of the cortisol block: 
```{r structure_2}
str(cortisol_block)
```
We see that the cortisol block contains 118 rows and 7 columns: 

1. **`StudyID`**: a patient identifier, which is a string, e.g. `cnt_101_bl`
2. **`M_S1`**, **`M_S2`**, ..., **`M_S6`**: the cortisol measure at each of six timepoints, which is a nonnegative number

We check for missing values:
```{r missings_2}
sapply(cortisol_block, function(x) sum(is.na(x)))
```

* There are no missing values.

### 1) SubjectID
No subject identifiers are duplicated: 
```{r dupes_2}
cortisol_block <- cortisol_block %>% 
  rename("SubjectID" = "StudyID")

table(duplicated(cortisol_block$SubjectID))
```

### 2) M_S1 ~ M_S6

```{r summary_2}
summary(cortisol_block)
```

Since the cortisol data are time-series, we will want to pivot the six timepoints into a single column for EDA.
```{r pivot_2}
cortisol_long <- cortisol_block %>%
  pivot_longer(!SubjectID, names_to="timepoint", values_to="cortisol") %>%
  mutate(timepoint = strtoi(str_sub(timepoint, 4)))   # extract timepoint and cast to integer
head(cortisol_long)
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save_2}
cortisol_wide <- cortisol_block
# save(cortisol_wide, cortisol_long, file = "./data/cortisol.RData")
```

## 3. FMRI data

```{r read_3}
fmri_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                         sheet = "ROI_MIST_block2")
```

We examine the structure of the FMRI block: 
```{r structure_3}
str(fmri_block)
```
We see that the FMRI block contains 928 rows and 3 columns: 

1. **`Subject`**: a patient identifier, which is a string, e.g. `cnt_101_bl`
2. **`ROI`**: Region of Interest
3. **`MeanPerSigChg`**: Mean per signal change

We check for missing values:
```{r missings_3}
sapply(fmri_block, function(x) sum(is.na(x)))
```

* There are no missing values.

### 1) SubjectID
There are total of 116 unique subjects.
```{r unique_3_1}
fmri_block <- fmri_block %>% 
  rename("SubjectID" = "Subject")

length(unique(fmri_block$SubjectID))
```

### 2) ROI
There are total of 8 unique ROIs.
```{r unique_3_2}
length(unique(fmri_block$ROI))
```

### 3) MeanPerSigChg

Since there are 116 unique subjects and 8 unique ROIs, we will need to reshape the dataset from long to wide format. The value in each cell of the resulting dataset will be the _MeanPerSigChg_ for a given subject and ROI combination. This transformation will allow us to more easily compare the signal changes across different brain regions and subjects.
```{r pivot_3}
fmri_wide <- fmri_block %>% 
  pivot_wider(names_from = ROI, values_from = MeanPerSigChg)
head(fmri_wide, 5)
```
After applying this transformation, the resulting dataset have 116 rows and 9 columns, where the first column is the subject ID and the remaining 8 columns represent the ROIs.

```{r summary_3}
summary(fmri_wide)
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save_3}
fmri_long <- fmri_block
# save(fmri_long, fmri_wide, file = "./data/fmri.RData")
```

## 4. Psych Profile data

```{r read_4}
psych_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                          sheet = "PsychProfile_block3")
```

We examine the structure of the Psych block: 
```{r structure_4}
str(psych_block)
```
We see that the Psych block contains 130 rows and 6 columns: 

1. **`SUBJ_ID`**: a patient identifier, which is a string, e.g. `cnt_101_bl`
2. **`STAITrait`**: 
3. **`STAIPost-Pre`**:
4. **`CogDisScore`**: score of cognition disorder 
5. **`Anhedonia`**:
6. **`CsopsNegative`**:

Before checking for missing values in the Pysch block, we need to convert _STAITrait_ from character format to numeric format.
```{r numeric_4}
psych_block <- psych_block %>% 
  mutate(STAITrait = as.numeric(na_if(STAITrait, "N/A")))
str(psych_block)
```

Now, we check for missing values:
```{r missings_4}
sapply(psych_block, function(x) sum(is.na(x)))
```

* There are no missing values in _SUBJ_ID_, _CogDisScore_, _Anhedonia_, and _AgeInMonths_.
* There are at least one missing value in _STAITrait_, and _STAIPost_.

### 1) SubjectID
No subject identifiers are duplicated: 
```{r dupes_4}
psych_block <- psych_block %>% 
  rename("SubjectID" = "SUBJ_ID")

table(duplicated(psych_block$SubjectID))
```

### 2 ~ 6) Psych measures
```{r summary_4}
summary(psych_block[, 2:6])
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save_4}
# save(psych_block, file = "./data/pysch.RData")
```

## 5. Working memory data

```{r read_5}
working_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                            sheet = "ROI_Nback_2vs0_block4")
```

We examine the structure of the Working memory block: 
```{r structure_5}
str(working_block)
```
We see that the Autonomic block contains 1024 rows and 4 columns: 

1. **`Subject`**: a patient identifier, which is a string, e.g. `cnt_101_bl`
2. **`ROI`**: Region of Interest
3. **`Status`**:
4. **`Mean`**:

We check for missing values:
```{r missings_5}
sapply(working_block, function(x) sum(is.na(x)))
```

* There are no missing values.

### 1) SubjectID
There are total of 8 unique subjects.
```{r unique_5_1}
working_block <- working_block %>% 
  rename("SubjectID" = "Subject")

length(unique(working_block$SubjectID))
```

### 2) ROI
There are total of 8 unique ROIs.
```{r unique_5_2}
length(unique(working_block$ROI))
```

### 3) Status
There are total of 8 unique ROIs.
```{r unique_5_3}
length(unique(working_block$Status))
```

Based on this, each of the 64 subjects has 16 (= 8 ROIs x 2 status levels) `Mean` values.

### 4) Mean
```{r summary_5}
summary(working_block$Mean)
```

Since there are 64 unique subjects and 8 unique ROIs with 2 status levels, we can reshape the dataset from wide to long format. The value in each cell of the resulting dataset will be the _Mean_ for a given combination of subject, ROI and the status. This transformation will allow us to more easily compare the signal changes across different brain regions and subjects with given status.
```{r pivot_5}
working_wide <- working_block %>% 
  pivot_wider(names_from = c("ROI", "Status"),
              values_from = "Mean") 
head(working_wide, 5)
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save_5}
working_long <- working_block
# save(working_long, working_wide, file = "./data/working.RData")
```

## 6. Autonomic data

```{r read_6}
autonomic_block <- read_excel("./data/cognit_stor765_data.xlsx", 
                              sheet = "HRV_SAA_block7")
```

We examine the structure of the Autonomic block: 
```{r structure_6}
str(autonomic_block)
```
We see that the Autonomic block contains 174 rows and 10 columns: 

1. **`Subject`**: a patient identifier, which is a string, e.g. `cnt_101_bl`
2. **`HRV_MIST1`**, **`HRV_MIST2`**, **`HRV_MIST3`**: heart-rate variability from 1 to 3
3. **`AA_S1`**, **`AA_S2`**, ..., **`AA_S6`**: salivary alpha amylase from 1 to 6

We check for missing values:
```{r missings_6}
sapply(autonomic_block, function(x) sum(is.na(x)))
```

* The columns **`HRV_MIST`** and **`AA_S`** have a high number of missing values, with approximately 50% of the elements in each column being missing.

### 1) SubjectID
No subject identifiers are duplicated: 
```{r dupes_6}
autonomic_block <- autonomic_block %>% 
  rename("SubjectID" = "Subject")

table(duplicated(autonomic_block$SubjectID))
```

### 2 ~ 6) Autonomic measures
```{r summary_6}
summary(autonomic_block[, 2:10])
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save_6}
# save(autonomic_block, file = "./data/autonomic.RData")
```