---
title: 'PCA: Principal Component Analysis'
author: "Amy Song"
date: "April 17, 2023"
output:
  html_document:
    df_print: paged
---

```{r, message = F}
source_rmd = function(file, ...) {
  tmp_file = tempfile(fileext = ".R")
  on.exit(unlink(tmp_file), add = TRUE)
  knitr::purl(file, output = tmp_file)
  source(file = tmp_file, ...)
}

source_rmd("../00_manage_packages.Rmd")
```
## Loading Data
```{r}
# Cortisol Data
load('../data/cortisol.RData')
# fMRI Data
load('../data/fmri.Rdata')
# Working Memory Data
load('../data/working.Rdata')
# demographics Data
load('../data/demographics.Rdata')
# psych data
load('../data/pysch.Rdata')
```

# Background: PCA and Scree Plot

Principal Components Analysis is used to find a lower dimensional structure from the high dimensional data. For data with correlated variables, principal components allow us to summarize the set of variables with a smaller number of representative variables that collectively explain most of the variability in the original set [2]. 

For each of the data block given, it is in the form of a matrix, $X \in \mathbb{R}^{n\times p}$, where $n$ is the number of observations, which for this project are the patients, and $p$ is the number of features, which differs for each of the data block for this project. When $p$ is large, a lower dimensional representation of the data matrix may capture a lot of information of the original data, reducing the... To capture this, "PCA finds a small number of dimensions that are interesting as possible, where the concept of *interesting* is measured by the amount that the observations vary along each dimension. Each of the dimensions found by PCA is a linear combination of the $p$ features" [2]

# PCA and Scree-Plot on data blocks
Using correlation matrix of each data block to standardize 
The number of components selected for each PCA of data block if 95% of the data is explained. 

## Cortisol (block 1)
Checking the dimension of the data block and whether there is `Subject` in the column. This column will be not be considered in the PCA. 
```{r}
dim(cortisol_wide)
names(cortisol_wide)
```

```{r, warning = F}
# PCA of the block: the matrix is scaled
cortisol_pca<- prcomp(cortisol_wide[,2:7], scale = F)
# Scree plot of the PCA
cumulated.var <- get_eig(cortisol_pca)$cumulative.variance.percent
cumulated.var.labs <- paste(round(cumulated.var,1),'%')
cumulated.var.labs[1] <- NA
factoextra::fviz_eig(cortisol_pca, addlabels = T, title = "Scree Plot for Cortisol")  + 
  geom_line(aes(y = cumulated.var), color = "lightslategrey") + 
  geom_point(aes(y = cumulated.var), color = "lightslategrey") +
  geom_text(aes(label = cumulated.var.labs, y = cumulated.var + 5), color = "lightslategrey")
```
```{r}
# Summary of the PCA
summary(cortisol_pca)
```
* As seen in the scree plot and the summary of the PCA, *3 Components* explains 99% of the variances in the data. The following is the biplot, which plots both scores and loading values of the first 2 principal components. The points on the plot indicate the scores of each observations and the blue errors indicate the loading vectors of the variables. For instance, the loading value for `M_S3` on the first PC is 0.2795931 and 0.3229215 for second PC, so the vector is centered at (0.2795931, 0.3229215)

```{r}
# Bi-Plot of the PCA: both scores and loading
autoplot(cortisol_pca, label = FALSE, loadings.label = TRUE, loadings.label.repel = TRUE) + 
  geom_vline(xintercept = 0, color = "gray28", lwd=0.25) + 
  geom_hline(yintercept = 0, color = "gray28", lwd=0.25) 
# loading values of PCA
cortisol_pca$rotation
```

```{r}
# Bar plot of the Contributions of each variables in the selected PC
factoextra::fviz_contrib(cortisol_pca,
                 choice = "var",
                 axes = 3)
```

## fMRI Block
Checking the dimension of the data block and whether there is `Subject` in the column. This column will be not be considered in the PCA. 
```{r}
dim(fmri_wide)
names(fmri_wide)
```

```{r, warning = F}
# PCA of the block: the matrix is centered
fmri_pca <- prcomp(fmri_wide[,2:9])
# Scree plot of the PCA
cumulated.var <- get_eig(fmri_pca)$cumulative.variance.percent
cumulated.var.labs <- paste(round(cumulated.var,1),'%')
cumulated.var.labs[1] <- NA
factoextra::fviz_eig(fmri_pca, addlabels = T, title = "Scree Plot for fMRI")  + 
  geom_line(aes(y = cumulated.var), color = "lightslategrey") + 
  geom_point(aes(y = cumulated.var), color = "lightslategrey") +
  geom_text(aes(label = cumulated.var.labs, y = cumulated.var + 5), color = "lightslategrey")
```
```{r}
# Summary of the PCA
summary(fmri_pca)
```
* As seen in the scree plot and the summary of the PCA, *5 Components* explains 95% of the variances in the data. The following is the biplot, which plots both scores and loading values of the first 2 principal components. The points on the plot indicate the scores of each observations and the blue errors indicate the loading vectors of the variables. For instance, the loading value for `amyg_R_stor` on the first PC is -0.4315 and 0.0300 for second PC, so the vector is centered at (-0.4315, 0.0300)

```{r}
# Bi-Plot of the PCA: both scores and loading
ggplot2::autoplot(fmri_pca, label = FALSE, loadings.label = TRUE, loadings.label.repel = T) + 
  geom_vline(xintercept = 0, color = "gray28", lwd=0.25) + 
  geom_hline(yintercept = 0, color = "gray28", lwd=0.25) 
# loading values of PCA
fmri_pca$rotation
```
```{r}
# Bar plot of the Contributions of each variables in the selected PC
factoextra::fviz_contrib(fmri_pca,
                 choice = "var",
                 axes = 5)
```

## Working Memory 

Checking the dimension of the data block and whether there is `Subject` in the column. 
```{r}
dim(working_wide)
names(working_wide)
```

```{r, warning = F}
# PCA of the block: the matrix is centered
#wm_corr <- cor(wm_df)
wm_pca <- prcomp(working_wide)
# Scree plot of the PCA
cumulated.var <- get_eig(wm_pca)$cumulative.variance.percent[1:10]
cumulated.var.labs <- paste(round(cumulated.var,1),'%')
cumulated.var.labs[1] <- NA
factoextra::fviz_eig(wm_pca, addlabels = T, title = "Scree Plot for Worming Memory")  + 
  geom_line(aes(y = cumulated.var), color = "lightslategrey") + 
  geom_point(aes(y = cumulated.var), color = "lightslategrey") +
  geom_text(aes(label = cumulated.var.labs, y = cumulated.var + 5), color = "lightslategrey")
```
```{r}
# Summary of the PCA
summary(wm_pca)
```

* As seen in the scree plot and the summary of the PCA, *9 Components* explains 95% of the variances in the data. The following is the biplot, which plots both scores and loading values of the first 2 principal components. The points on the plot indicate the scores of each observations and the blue errors indicate the loading vectors of the variables. For instance, the loading value for `amyg_L_stor_PostStress` on the first PC is 0.039938048 and -0.174290802 for second PC, so the vector is centered at (0.039938048, -0.174290802)

```{r}
# Bi-Plot of the PCA: both scores and loading
ggplot2::autoplot(wm_pca, label = FALSE, loadings.label = TRUE, loadings.label.repel = T) + 
  geom_vline(xintercept = 0, color = "gray28", lwd=0.25) + 
  geom_hline(yintercept = 0, color = "gray28", lwd=0.25) 
# loading values of PCA
wm_pca$rotation[,c(1,2,9)]
```

## Demographics 

```{r}
dim(demographics_block)
names(demographics_block)
```

```{r, warning=F}
# PCA of the block: the matrix is centered and scaled
demo_pca <- demographics_block %>% mutate(Sex = as.numeric(Sex)) %>% select(-c(SubjectID,AgeInYears)) %>% drop_na(.) %>% prcomp(scale = T)
# Scree plot of PCA with cumulated sum of percentage of variance explained
cumulated.var <- get_eig(demo_pca)$cumulative.variance.percent
cumulated.var.labs <- paste(round(cumulated.var,1),'%')
cumulated.var.labs[1] <- NA
factoextra::fviz_eig(demo_pca, addlabels = T, title = "Scree Plot for Demographics")  + 
  geom_line(aes(y = cumulated.var), color = "lightslategrey") + 
  geom_point(aes(y = cumulated.var), color = "lightslategrey") +
  geom_text(aes(label = cumulated.var.labs, y = cumulated.var + 5), color = "lightslategrey")
```
```{r}
summary(demo_pca)
```
* As seen in the scree plot and the summary of the PCA, *4 Components* explains at least 95% of the variances in the data. The following is the biplot, which plots both scores and loading values of the first 2 principal components.

```{r}
# Bi-Plot of the PCA: both scores and loading
ggplot2::autoplot(demo_pca, label = FALSE, loadings.label = TRUE, loadings.label.repel = T, data = demographics_block %>% select(-c(SubjectID,AgeInYears)) %>% drop_na(.), colour = 'Sex')  + 
  geom_vline(xintercept = 0, color = "gray28", lwd=0.25) + 
  geom_hline(yintercept = 0, color = "gray28", lwd=0.25) 
# loading values of PCA
demo_pca$rotation
# color bi-plot
```

```{r}
# Bar plot of the Contributions of each variables in the selected PC
factoextra::fviz_contrib(demo_pca,
                 choice = "var",
                 axes = 4)
```

## Psych Profile
drop the rows with missing values
```{r}
dim(psych_block)
names(psych_block)
```

```{r, warning=F}
psych_pca <- psych_block %>% select(-SubjectID) %>% drop_na(.) %>% prcomp(scale = T)
# Scree plot of PCA with cumulated sum of percentage of variance explained
cumulated.var <- get_eig(psych_pca)$cumulative.variance.percent
cumulated.var.labs <- paste(round(cumulated.var,1),'%')
cumulated.var.labs[1] <- NA
factoextra::fviz_eig(psych_pca, addlabels = T, title = "Scree Plot for Psych Profile")  + 
  geom_line(aes(y = cumulated.var), color = "lightslategrey") + 
  geom_point(aes(y = cumulated.var), color = "lightslategrey") +
  geom_text(aes(label = cumulated.var.labs, y = cumulated.var + 5), color = "lightslategrey")
```

```{r}
summary(psych_pca)
```

* As seen in the scree plot and the summary of the PCA, *4 Components* explains 95% of the variances in the data. The following is the biplot, which plots both scores and loading values of the first 2 principal components.

```{r}
# Bi-Plot of the PCA: both scores and loading
ggplot2::autoplot(psych_pca, label = FALSE, loadings.label = TRUE, loadings.label.repel = T) + 
  geom_vline(xintercept = 0, color = "gray28", lwd=0.25) + 
  geom_hline(yintercept = 0, color = "gray28", lwd=0.25) 
# loading values of PCA
psych_pca$rotation[,c(1,2,4)]
```

```{r}
# Bar plot of the Contributions of each variables in the selected PC
factoextra::fviz_contrib(psych_pca,
                 choice = "var",
                 axes = 4)
```