---
title: "01_make_data"
author: "KyungjinSohn"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, include=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
} 
p_load(readxl,
       dplyr)
```

## Demographics data
```{r df_read}
demographics_df <- read_excel("./source_data/cognit_stor765_data.xlsx", sheet = "Demographics_block5")
```

The dataset includes demographic information for 130 subjects and consists of 6 variables.
```{r df_dim}
dim(demographics_df)
```

We have 6 different types of variables for each subject: Subject ID, sex, age in months, own puberty level, parent's puberty level, and average parent education.
```{r df_str}
str(demographics_df)
```

### 1) ID
No subject IDs are duplicated. 
```{r}
# 1. Subject ID
table(duplicated(demographics_df$Subject))
```

### 2) Sex
Next, we changed the Sex variable to a factor type variable.
```{r}
# 2. Sex
demographics_df <- demographics_df %>% 
     mutate(Sex = factor(Sex, levels = c("M", "F")))
str(demographics_df$Sex)
```

### 3) Age In Months
Since we only had the age in months, we added a new column to show the age in years rounded to the second decimal place.
```{r}
# 3. Age in Months
demographics_df <- demographics_df %>% 
     mutate(AgeInYears = round(AgeInMonths/12, 2)) %>% 
     relocate(AgeInYears, .after = AgeInMonths)
```

### 4 ~ 6) Own puberty level, Parent's puberty level, and Average parent education
The own puberty level variable ranges from 1 to 5, with 29 missing values. The parent's puberty level variable also ranges from 1 to 5, with 21 missing values. Lastly, the average parent education variable ranges from 1 to 9, with 6 missing values.
```{r df_sum}
summary(demographics_df)
```

After making the modifications mentioned above, we saved the updated data frame.
```{r save}
save(demographics_df, file = "./demographics_df.RData")
```