---
title: "02_EDA"
author: "KyungjinSohn"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, include=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
} 
p_load(naniar,
       ggplot2,
       tidyr,
       dplyr,
       GGally)
```

```{r read}
load("./demographics_df.RData")
```

## Demographics data

### 1. Missing data

There are missing values in the Own puberty level, Parent's puberty level, and Average parent education variables. Specifically, 22% of the observations have missing values for Own puberty level, 16% are missing for Parent's puberty level, and 5% of the Average parent education variable is missing.

```{r heatmap, fig.align='center', fig.width=9, fig.height=7}
vis_miss(demographics_df) + 
     theme(axis.text.x = element_text(angle = 75)) +
     labs(title = "Heatmap of missing data in the demographics")
```

To be specific, 29 observations for Own puberty level, 21 observations for Parent's puberty level, and 6 observations for Average parent education variable are missing.
```{r}
sapply(demographics_df, function(x) sum(is.na(x)))
```

One subject (`cnt_277_bl`) has missing values for all three variables.
```{r na3}
allNA <- which(apply(demographics_df, MARGIN = 1, function(x) sum(is.na(x))) == 3)
demographics_df[allNA,]
```

Additionally, there are 9 subjects who have two missing values among the three columns. Of these 9 subjects, 6 have two missing values in the two puberty level columns, and 3 have two missing values in both the Parental puberty level and Parental education average columns.
```{r na1}
twoNA <- which(apply(demographics_df, MARGIN = 1, function(x) sum(is.na(x))) == 2)
demographics_df[twoNA,]
```

If we remove observations with at least one missing value, there are 85 subjects remaining.
```{r}
nrow(demographics_df %>% drop_na(.))
```

### 2. Associations among variables

* There are more male subjects than female.
* The Age, Own puberty level, and Parent's puberty level variables are fairly spread out.
* The Average parent education level variable is left-skewed, with most values ranging between 6 to 10.
* Among the dataset, females have a higher average own puberty level than males.
* There seems to be a strong positive correlation between Age in years and Parent's puberty level.
* Additionally, there is a high correlation between Own puberty level and Parent's puberty level.

```{r scatter, warning=FALSE, message=FALSE, fig.align='center', fig.width=9}
ggpairs(demographics_df, columns = c(2, 4:7), 
        title = "Scatterplot matrix (pairs plots) of variables")
```
