---
title: "01_make_data"
author: "Tom Keefe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, include=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
} 
p_load(readxl,
       dplyr)
```

## Cortisol data
```{r df_read}
cortisol_block <- read_excel("./source_data/cognit_stor765_data.xlsx", 
                              sheet = "Cortisol_MRI_block1")
```

We examine the structure of the cortisol block using `str`: 

```{r df_str}
str(cortisol_block)
```
We see that cortisol block contains 118 rows and 7 columns: 

* a patient identifier, _StudyID_, which is a string, e.g. `cnt_101_bl`
* the cortisol measure at each of six timepoints, _M_S1_, _M_S2_, ..., _M_S6_, which is a nonnegative number.

We check that no patient identifiers are duplicated:
```{r}
table(duplicated(cortisol_block$StudyID))
```

Indeed, none of the 118 values show up more than once. The data should now be ready for EDA, so we save the cortisol block with the following:

```{r save}
save(cortisol_block, file = "./cortisol_block.RData")
```
