---
title: "01_make_data"
author: "Tom Keefe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
} 
p_load(readxl,
       dplyr,
       tidyr,
       stringr,
       ggplot2)
```



## Cortisol data
```{r read}
cortisol_block <- read_excel("./source_data/cognit_stor765_data.xlsx", 
                              sheet = "Cortisol_MRI_block1")
```

We examine the structure of the cortisol block using `str`: 

```{r structure}
str(cortisol_block)
```
We see that the cortisol block contains 118 rows and 7 columns: 

* a patient identifier, _StudyID_, which is a string, e.g. `cnt_101_bl`
* the cortisol measure at each of six timepoints, _M_S1_, _M_S2_, ..., _M_S6_, which is a nonnegative number. 

We check that no patient identifiers are duplicated:
```{r dupes}
table(duplicated(cortisol_block$StudyID))
```

Indeed, none of the 118 values show up more than once.

We check for missing values:

```{r missings}
sum(is.na(cortisol_block))
```
There are no missing values.


Since the cortisol data are time-series, we will want to pivot the six timepoints into a single column for EDA.

```{r pivot}
cortisol_long = cortisol_block %>%
    pivot_longer(!StudyID, names_to="timepoint", values_to="cortisol") %>%
    mutate(timepoint = strtoi(str_sub(timepoint, 4)))   # extract timepoint and cast to integer
head(cortisol_long)
```

```{r}
save(cortisol_block, cortisol_long, file = "./cortisol.RData")
```










