---
title: "02_EDA"
author: "Tom Keefe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, include=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
} 
p_load(ggplot2,
       tidyr,
       dplyr,
       GGally)
```

```{r read}
load("./cortisol.RData")
```

## EDA of cortisol data

### 1. Missing data

There is no missing data in the cortisol block:
```{r}
sum(is.na(cortisol_block))
```

We now move on to visualizing the marginal and joint distributions of the cortisol variables.

### 2. Associations among variables

```{r scatter, warning=FALSE, message=FALSE, fig.align='center', fig.width=9}
ggpairs(cortisol_block, columns = c(2:7), 
        title = "Scatterplot matrix (pairs plots) of variables")
```



## Boxplots

```{r}
cortisol_long %>% ggplot(aes(y=cortisol)) +
    facet_grid(~timepoint) + 
    geom_boxplot() + 
    theme(axis.text.x = element_blank()) + 
    ggtitle("Distribution of cortisol level, facetted by timepoint")
```

We can see that the marginal distributions are right-skewed, and each the marginal distribution of each timepoint has several outliers.
These observations suggest that a log transformation may be helpful for downstream analysis.

Next, we will use a line plot to determine belong to a small group of subjects. In order to prevent this from being a "spaghetti plot," we limit the vertical axis to values above 0.25, which is where the outliers begin to occur.

```{r spag}
cortisol_long %>% 
    filter(cortisol > 0.25) %>%
    ggplot(aes(y=cortisol)) + 
    geom_line(aes(x=timepoint, group=StudyID, col=StudyID), show.legend=F) +
    ggtitle("Spaghetti plot of cortisol readings above 0.25") + 
    ylim(c(0.25, NA))
```

We can see that the outlier points in the boxplots refer to a small number of curves. 

## PCA

```{r pca}
pca = cortisol_block %>%
    select(-c(StudyID)) %>%
    prcomp()
```

```{r}
summary(pca)
```

```{r}
plot(pca)
```




