---
title: "01 Make Data: Block 2"
author: "Amy Song"
date: "02/02/2023"
output: html_notebook
---

Setting up libraries that will be used in reading the data.
```{r}
library(dplyr)
library(tidyverse)
```

Reading the data from the raw dataset from the client. I will be reading data on signals from fMRI on 4 ROI, regions of interest (sheet name: "ROI_MIST_block2") and correlation of these signals from fMRI ROIs and the status of stress (either pre-stress or post-stress) (sheet name: "ROI_Nback_2vs0_block4"). 
```{r}
fmri <- readxl::read_excel('cognit_stor765_data.xlsx', sheet = "ROI_MIST_block2")
fmri.corr <- readxl::read_excel('cognit_stor765_data.xlsx', sheet = "ROI_Nback_2vs0_block4")
```

The dimension of the fmri data is shown below: 928 rows and 3 columns, where the names of the columns are the following: "Subject","ROI","MeanPerSigChg". And there are total of 116 unique subjects in the data with 8 unique ROIs. 
```{r}
dim(fmri)
```
```{r}
names(fmri)
```
```{r}
length(unique(fmri$Subject))
```

As the dimensions of the data show, transformation on the data needs to be made. The following transformation of the data leads to the data to have a dimension of 116 rows and 9 columns. 
```{r}
fmri_block <- fmri %>% 
  pivot_wider(names_from = ROI, values_from = MeanPerSigChg)
head(fmri_block)
dim(fmri_block)
```
The transformed data will be saved as a `Rdata` formt in Rdata folder. 
```{r}
save(fmri_block, file = "./Rdata/fmri_block.Rdata")
```


The correlation matrix of 
```{r}
dim(fmri.corr)
```

