---
title: "02_EDA_Block_3_7"
author: "SooHyun Kim"
date: "2023-04-07"
output: html_document
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, include=FALSE}
library(dplyr)
library(readxl)
library(naniar)
library(UpSetR)
library(ggplot2)
library(GGally)
```

# Load Data
From here we will only look at the Prostate Data
```{r}
load("block3_psych_df.RData")
load("block7_hrv_saa_df.RData")
```

# Block 3 Psych PRofile
## Missing Data
### Percentage of Missing Data
```{r}
vis_miss(block3_psych_df)
```
From the plot above, we can see that there is only missing variables for the `STAIPost-Pre` variable of 4%, which 5 out of 130 observations. 

```{r}
block3_psych_df[is.na(block3_psych_df$`STAIPost-Pre`),]
```
### Exploring the Missing Data Points
```{r}
# using  geom_miss_point()
ggplot(block3_psych_df,
       aes(x = STAITrait,
           y = `STAIPost-Pre`)) +
 geom_miss_point()

# using  geom_miss_point()
ggplot(block3_psych_df,
       aes(x = CogDisScore,
           y = `STAIPost-Pre`)) +
 geom_miss_point()

# using  geom_miss_point()
ggplot(block3_psych_df,
       aes(x = Anhedonia,
           y = `STAIPost-Pre`)) +
 geom_miss_point()

# using  geom_miss_point()
ggplot(block3_psych_df,
       aes(x = Anhedonia,
           y = `STAIPost-Pre`)) +
 geom_miss_point()
```
Does not seem to be a clear trend in the missing data points.

```{r}
gg_miss_var(block3_psych_df, show_pct = TRUE)
```

## Scatterplot Matrix (Pairs Plots)
```{r}
block3_psych_df[, 2:6] <- lapply(block3_psych_df[, 2:6], as.numeric)
ggpairs(block3_psych_df[,2:6])

ggcorr(block3_psych_df[,2:6])

```




## PCA scatterplot matrix and Scree Plots

# Block 7 HRV SAA
## Missing Data
### Percentage of Missing Data
```{r}
vis_miss(block7_hrv_saa_df[,1:10])
```

Missing data points is a significant issue in Block 7. Missing data appears in $47.3%$ of the data. The `SAA` variables seem to have a high number of missing values with over $50%$ for all `AA` variables. For the `HRV_MIST` variables have a missing rate all over $40%$

### Exploring the Patterns of Missing Data
```{r}
gg_miss_upset(block7_hrv_saa_df[,2:10], nsets = n_var_miss(block7_hrv_saa_df))

```
From the full missing data, 53 data points do not have any data inputs. Just considering the `SAA` variables 26 data points do not have any information about `SAA`. For `HRV` there are 7 data points.

```{r}
gg_miss_upset(block7_hrv_saa_df[,2:4])
```
Out of the `HRV`, `HRV_MIST3` has the most missing values. There are 68 cases where all 3 variables are missing, 7 of `HRV_MIST2` and `HRV_MIST3` missing, 2 of `HRV_MIST1` and `HRV_MIST2`, and 1 of `HRV_MIST1` and `HRV_MIST3`.

```{r}
gg_miss_upset(block7_hrv_saa_df[,5:10])
```
There are 87 instances of all `SAA` variables missing. Other combinations of missing data are relatively smaller compared to data sets with all missing points.


### Exploring the Missing Data Points
```{r}
# using  geom_miss_point()
ggplot(block7_hrv_saa_df,
       aes(x = HRV_MIST1,
           y = HRV_MIST2)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = HRV_MIST1,
           y = HRV_MIST3)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = HRV_MIST2,
           y = HRV_MIST3)) +
 geom_miss_point()
```
```{r}
# using  geom_miss_point()
ggplot(block7_hrv_saa_df,
       aes(x = AA_S1,
           y = AA_S2)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S1,
           y = AA_S3)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S1,
           y = AA_S4)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S1,
           y = AA_S5)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S1,
           y = AA_S6)) +
 geom_miss_point()

##2
ggplot(block7_hrv_saa_df,
       aes(x = AA_S2,
           y = AA_S3)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S2,
           y = AA_S4)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S2,
           y = AA_S5)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S2,
           y = AA_S6)) +
 geom_miss_point()

##3
ggplot(block7_hrv_saa_df,
       aes(x = AA_S3,
           y = AA_S4)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S3,
           y = AA_S5)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S3,
           y = AA_S6)) +
 geom_miss_point()

## 4
ggplot(block7_hrv_saa_df,
       aes(x = AA_S4,
           y = AA_S5)) +
 geom_miss_point()

ggplot(block7_hrv_saa_df,
       aes(x = AA_S4,
           y = AA_S6)) +
 geom_miss_point()

##5
ggplot(block7_hrv_saa_df,
       aes(x = AA_S5,
           y = AA_S6)) +
 geom_miss_point()
```

```{r}
gg_miss_var(block7_hrv_saa_df, show_pct = TRUE)
```

## Scatterplot Matrix (Pairs Plots)
```{r}
block7_hrv_saa_df[, 2:10] <- lapply(block7_hrv_saa_df[, 2:10], as.numeric)
ggpairs(block7_hrv_saa_df[,2:10])
ggcorr(block7_hrv_saa_df[,2:10])

```

```{r}
ggpairs(block7_hrv_saa_df[,2:4])
ggcorr(block7_hrv_saa_df[,2:4])
```

```{r}
ggpairs(block7_hrv_saa_df[,5:10])
ggcorr(block7_hrv_saa_df[,5:10]) 
```


