---
title: "03_MissingData"
author: "SooHyun Kim"
date: "2023-04-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r lib, include=FALSE}
if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
} 
p_load(readxl,
       naniar,
       ggplot2,
       tidyr,
       dplyr,
       GGally)
```

# Load File
```{r}
setwd("C:/Users/SooHyun Sam Kim/OneDrive/문서/GitHub/STOR765_Spring2023_TeamDS")

# cortisol (1)
cortisol_df <- read_excel("./kyungjin/source_data/cognit_stor765_data.xlsx", sheet = "Cortisol_MRI_block1")

# fmri (2,5)
load("./amy/Rdata/fmri_block.RData")

# demographic (4)
load("./kyungjin/demographics_df.RData")

# psych (3)
load("./soohyun/block3_psych_df.RData")

# hrv_saa (6)
load("./soohyun/block7_hrv_saa_df.RData")

# working memory (7)
load("./taebin/RData/wm_df.RData")
```

## Block 1. Cortisol MRI
### A. Missing Data
* No missing data for block 1
```{r}
nrow(cortisol_df[complete.cases(cortisol_df),])/nrow(cortisol_df)
```

### B. Summary of variables
```{r}
summary(cortisol_df)
```


### C. Associations among variables

* All regions are significantly correlated (all over $0.3$)
* High correlation between `M_S1`,`M_S2`, and `M_S3` all with a correlation above $0.6$
* `M_S4`,`M_S5`, and `M_S6` are highly correlated


```{r scatter, warning=FALSE, message=FALSE, fig.align='center', fig.width=9}
ggpairs(cortisol_df, columns = c(2:7), 
        title = "Scatterplot matrix (pairs plots) of variables")
```

## Block 2. FMRI

### A. Missing Data
* No missing data for block 2
```{r}
nrow(fmri_block[complete.cases(fmri_block),])/nrow(fmri_block)
```

### B. Summary of variables
```{r}
summary(fmri_block)
```


### C. Associations among variables
### Change
* All regions are significantly correlated (all over $0.3$)
* High correlation between `M_S1`,`M_S2`, and `M_S3` all with a correlation above $0.6$
* `M_S4`,`M_S5`, and `M_S6` are highly correlated


```{r scatter, warning=FALSE, message=FALSE, fig.align='center', fig.width=9}
ggpairs(fmri_block, columns = c(2:9), 
        title = "Scatterplot matrix (pairs plots) of variables")
```

## Block 3. Psych Profile Block
### A. Missing Data
* Out of 130 observations 8 rows have at least one missing input.
```{r}
nrow(block3_psych_df[complete.cases(block3_psych_df),])/nrow(block3_psych_df)
nrow(block3_psych_df) - nrow(block3_psych_df[complete.cases(block3_psych_df),])
```
```{r psychheatmap, fig.align='center', fig.width=9, fig.height=7}
vis_miss(block3_psych_df) + 
     theme(axis.text.x = element_text(angle = 75)) +
     labs(title = "Heatmap of Missing Data in Block 3. Psych Profile")
```
```{r}
sapply(block3_psych_df, function(x) sum(is.na(x)))
```

From the plot above, we can see that for variable `STAIPost-Pre` variable of 4%, which 5 out of 130 observations. We can also check that `STAITrait` has 4 missing values. One observation has both variables with missing. 

#### Exploring the Missing Data Points
```{r}
gg_miss_upset(block3_psych_df[,2:6], nsets = n_var_miss(block3_psych_df))
```

```{r}
# using  geom_miss_point()
ggplot(block3_psych_df,
       aes(x = STAITrait,
           y = `STAIPost-Pre`)) +
 geom_miss_point()

# using  geom_miss_point()
ggplot(block3_psych_df,
       aes(x = CogDisScore,
           y = `STAIPost-Pre`)) +
 geom_miss_point()

# using  geom_miss_point()
ggplot(block3_psych_df,
       aes(x = Anhedonia,
           y = `STAIPost-Pre`)) +
 geom_miss_point()

# using  geom_miss_point()
ggplot(block3_psych_df,
       aes(x = Anhedonia,
           y = `STAIPost-Pre`)) +
 geom_miss_point()
```

If we remove observations with at least one missing value, there are 122 subjects remaining.
```{r}
nrow(block3_psych_df %>% drop_na(.))
block3_psych_df2 <- block3_psych_df %>%  drop_na(.)
```

### B. Summary of variables
```{r}
summary(block3_psych_df)
```


### C. Associations among variables
### Change
* All regions are significantly correlated (all over $0.3$)
* High correlation between `M_S1`,`M_S2`, and `M_S3` all with a correlation above $0.6$
* `M_S4`,`M_S5`, and `M_S6` are highly correlated


```{r scatter, warning=FALSE, message=FALSE, fig.align='center', fig.width=9}
ggpairs(block3_psych_df, columns = c(2:6), 
        title = "Scatterplot matrix (pairs plots) of variables")
```


## Block 4. Demographic
### A. Missing Data
* Out of 130 observations 45 rows have at least one missing input.
```{r}
nrow(demographics_df[complete.cases(demographics_df),])/nrow(demographics_df)
nrow(demographics_df) - nrow(demographics_df[complete.cases(demographics_df),])
```

```{r heatmap, fig.align='center', fig.width=9, fig.height=7}
vis_miss(demographics_df) + 
     theme(axis.text.x = element_text(angle = 75)) +
     labs(title = "Heatmap of Missing Data in the Demographics")
```

To be specific, 29 observations for Own puberty level, 21 observations for Parent's puberty level, and 6 observations for Average parent education variable are missing.
```{r}
sapply(demographics_df, function(x) sum(is.na(x)))
```

One subject (`cnt_277_bl`) has missing values for all three variables.
```{r na3}
allNA <- which(apply(demographics_df, MARGIN = 1, function(x) sum(is.na(x))) == 3)
demographics_df[allNA,]
```

Additionally, there are 9 subjects who have two missing values among the three columns. Of these 9 subjects, 6 have two missing values in the two puberty level columns, and 3 have two missing values in both the Parental puberty level and Parental education average columns.
```{r na1}
twoNA <- which(apply(demographics_df, MARGIN = 1, function(x) sum(is.na(x))) == 2)
demographics_df[twoNA,]
```

#### Exploring the Missing Data Points

```{r}
gg_miss_upset(demographics_df[,2:7], nsets = n_var_miss(demographics_df))
```

```{r}
# using  geom_miss_point()
ggplot(demographics_df,
       aes(x = PubertyTannerSelf,
           y = PubertyTannerParent)) +
 geom_miss_point()

# using  geom_miss_point()
ggplot(demographics_df,
       aes(x = PubertyTannerSelf,
           y = ParentalEducationAvg)) +
 geom_miss_point()

# using  geom_miss_point()
ggplot(demographics_df,
       aes(x = PubertyTannerSelf,
           y = ParentalEducationAvg)) +
 geom_miss_point()
```

If we remove observations with at least one missing value, there are 85 subjects remaining.
```{r}
nrow(demographics_df %>% drop_na(.))
```

### B. Summary of variables
```{r}
summary(demographics_df)
```


### C. Associations among variables

* There are more male subjects than female.
* The Age, Own puberty level, and Parent's puberty level variables are fairly spread out.
* The Average parent education level variable is left-skewed, with most values ranging between 6 to 10.
* Among the dataset, females have a higher average own puberty level than males.
* There seems to be a strong positive correlation between Age in years and Parent's puberty level.
* Additionally, there is a high correlation between Own puberty level and Parent's puberty level.

```{r scatter, warning=FALSE, message=FALSE, fig.align='center', fig.width=9}
ggpairs(demographics_df, columns = c(2, 4:7), 
        title = "Scatterplot matrix (pairs plots) of variables")
```

# Block 5. Heart Rate Variability and SAA
### A. Missing Data
* Out of 174 observations 137 rows have at least one missing input.
```{r}
nrow(block7_hrv_saa_df[complete.cases(block7_hrv_saa_df),])/nrow(block7_hrv_saa_df)
nrow(block7_hrv_saa_df) - nrow(block7_hrv_saa_df[complete.cases(block7_hrv_saa_df),])
```
```{r heatmap, fig.align='center', fig.width=9, fig.height=7}
vis_miss(block7_hrv_saa_df) + 
     theme(axis.text.x = element_text(angle = 75)) +
     labs(title = "Heatmap of Missing Data in the HRV and SAA")
```

Missing data points is a significant issue in Block 7. Missing data appears in $47.3%$ of the data. The `SAA` variables seem to have a high number of missing values with over $50%$ for all `AA` variables. For the `HRV_MIST` variables have a missing rate all over $40%$

### Exploring the Patterns of Missing Data
```{r}
gg_miss_upset(block7_hrv_saa_df[,2:10], nsets = n_var_miss(block7_hrv_saa_df))
```

53 observations have all missing inputs. We will drop this block from analysis due to missing values issue.



### B. Summary of variables
```{r}
summary(block7_hrv_saa_df)
```



### C. Associations among variables

* The 3 measurements of Heart Rate Varaibility (HRV) is highly correlated with each other.
* All 6 measurements of SAA are highly correlated with each other.

```{r scatter, warning=FALSE, message=FALSE, fig.align='center', fig.width=9}
ggpairs(block7_hrv_saa_df, columns = c(2:10), 
        title = "Scatterplot matrix (pairs plots) of variables")
```


# Block 6. Working Memory
### A. Missing Data
* No missing data for block 6 Working memory.
```{r}
nrow(wm_df[complete.cases(wm_df),])/nrow(wm_df)
nrow(wm_df) - nrow(wm_df[complete.cases(wm_df),])
```


### B. Summary of variables
```{r}
summary(wm_df)
```


### C. Associations among variables

* The 3 measurements of Heart Rate Variability (HRV) is highly correlated with each other.
* All 6 measurements of SAA are highly correlated with each other.

```{r scatter, warning=FALSE, message=FALSE, fig.align='center', fig.width=9}
ggpairs(wm_df, columns = c(2:16), 
        title = "Scatterplot matrix (pairs plots) of variables")


ggcorr(wm_df[,2:16]) 
```
